Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-06-10 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix

\keywords{Multicriterial problems \and Global optimization \and Logistic regression}
\end{abstract}
%
%
%
\section{Introduction}
Artificial intelligence models formulated as decision-making problems are extremely diverse and attract significant attention of researchers due to the richness of theoretical background and practical importance. The high complexity of decision-making models, along with the use of existing effective approaches to their analysis, requires the development of new ideas that overcome these difficulties. A promising way to improve decision-making algorithms is to combine in their development the advantages of different approaches. For example, machine learning procedures effectively use built-in optimization methods to enhance the quality of learning.

@@

For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions and number of criteria. For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization. The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion. The results of the experiment have demonstrated significant advantages of the new algorithm over other methods having involved in comparison.

>>>>

\keywords{Multicriterial problems \and Global optimization \and Logistic regression}
\end{abstract}
%
%
%
\section{Introduction}
Artificial intelligence models formulated as decision-making problems are extremely diverse and attract significant attention of researchers due to the richness of theoretical background and practical importance. The high complexity of decision-making models requires the development of new ideas that reduce this complexity. A promising way to improve decision-making algorithms is to combine in their development the advantages of different approaches. For example, machine learning procedures effectively use built-in optimization methods to enhance the quality of learning.

@@

For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions. For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization. The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion. The results of the experiment have demonstrated significant advantages of the new algorithm over other methods having involved in comparison.
